---
title: "ndv24_HW_04"
author: "Nicholas Velotta | ndv24@uw.edu"
date: '`r format(Sys.time(), "%Y-%m-%d %H:%M")`'
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
output: 
    bookdown::html_document2:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_cap: yes
        keep_tex: yes
urlcolor: blue
---

```{r echo = FALSE, results = 'hide', message = FALSE, warning = FALSE, include = FALSE}
knitr::opts_chunk$set
library(tidyverse)
library(pander)
library(table1)
library(stargazer)
library(ggplot2)
library(knitr)
library(kableExtra)
library(magrittr)
library(psych)
library(readstata13)
library(animation)
library(captioner)
library(dplyr)
library(Rmisc)
```


# Questions about graph

```{r, echo=FALSE}
#creating data
gradpain <- c(rep(0,11), rep(1,1), rep(2,2), rep(3,6), rep(4,8), rep(5,10),
 rep(6,8), rep(7,30), rep(8,10), rep(9,6), rep(10,2))

#plotting
barplot(table(gradpain), las=1, ylab="Frequency", xlab="Pain Scale",
 main="How painful is graduate school?")

```

> How many graduate students are in the sample? Use R code to determine this.

```{r}
length(gradpain)
```
The sample size is `r length(gradpain)`

> What is the sample mean?

```{r}
mean(gradpain)
```
The mean of `gradpain` is `r round(mean(gradpain), 2)`

# Box 1 Questions
>Create a function, with these arguments:
1. the vector of data: "d.vec"
2. the size of the sample: "n"
The function will sample with replacement a sample of size "n" from the vector "d.vec". The
function will return a list that contains
1. the size of the sample
2. the mean of the sample

I was a bit confused as to what you're looking for here so I went with 2 interpretations:

This code (`newfunc`) produces a vector as part of the *body*

```{r}
newfunc <- function(x) {
    d.vec <- 1:10
    n <- sample(x = d.vec, size = 100, replace = TRUE)
    return(c( "average" = mean(n), "sample size" = length(n)))}

newfunc(1)
```

This code (`newfunc2`) asks for a vector as an argument and a sample size as part of the *arguments*
```{r}
newfunc2 <- function(d.vec, n) {
  s <- sample(x = d.vec, size = n, replace = TRUE)
  return(c( "average" = mean(s), "sample size" = length(s)))}
newfunc2(1:10, 150)
```

# Box 2 Questions 

>Use set.seed(7) then run your function passing in the "gradpain" vector calculated above and a
sample size of length(gradpain). Use a loop to do this 100 times and store all 100 returned
means.

```{r}
# initialize a vector
F <- NULL

# run the bootstrap
for (i in seq(from = 1, to = 100, by = 1)){
    # sample once
    s <- sample(x = gradpain, size = 100, replace = TRUE)
    # concatenate the result with the running result
    F <- c(F, mean(gradpain))
}
# mean and standard deviation of the bootstrap
mean(F)

```
> What is the mean of these 100 means?

The mean is `r mean(F)`. 

> What is the standard deviation of these 100 means?

The sd is `r sd(F)`.

# Box 3 Questions

> Write another function that performs the steps listed in Box 2. That should be a function with
these arguments:
1. the vector of data: "d.vec"
2. the size of the sample: "n"
3. the number of samples: "num.samples"
The function should sample with replacement a sample of size "n" from the vector "d.vec" and
does this "num.samples" times.
The function should return a list that contains
1. the size of each sample
2. the number of samples
3. a vector of length num.samples with the mean of each sample
4. the mean of the means
5. the standard deviation of the means
6. the 95% confidence interval around the mean
Run your function with the three arguments
d.vec = gradpain, n = length(gradpain), num.samples = 100

```{r}
T <- NULL

newfunc3 <- function(d.vec, n, num.samples) {
  for (i in seq(from = 1, to = num.samples, by = 1))
  s <- sample(x = d.vec, size = n, replace = TRUE)
  ci_95 <- Rmisc::CI(x = d.vec, ci = 0.95)
  return(c( "mean of means" = mean(s), "sd of means" = sd(s), "sample size" = length(s), "number of samples" = num.samples, "CI" = ci_95)) 
  T <- c(T, mean(d.vec), num.samples)} #my vector wont generate for this one...any ideas?
answers <-newfunc3(gradpain, length(gradpain), 100)

pander(answers)
```

> What does your function return for the mean of means?

5.681

> What does your function return for the standard deviation of means?

2.587  

> What does your function return for the 95% confidence interval around the mean?

upper limit of 6.054 and lower limit of 4.946
